### 2025-05-05 chapters/2-particles.tex
Pre-correction version of thesis

Rather than scattering rocks, we will be concerned with the scattering of subatomic particles.
%
First, what is a particle?
%
To answer this, let us point out that an electron moving to your left is a different state than an electron moving to your right;
%
furthermore, if you say that it is moving to the left, someone facing you will say that it is moving to the right!
%
But you should both agree, regardless of the direction of its motion or spin, that it is the same \textit{type} of particle:
%
an electron.
%
So a particle should be a \textit{set} or \textit{collection} of states which different observers agree correspond to the same particle.

\begin{definitionbox}{Particle}{particle}
    A particle is a \textit{set} of states which is the same for every (inertial) observer.

    A \vocab{particle} is a \textit{set} of quantum states characterized by mass, spin, and charge -- their \textit{quantum numbers} -- such that the overall set of states remains unchanged under Poincar\'e transformations:
    %
    translations in space/time, rotations, and boosts.
\end{definitionbox}

Some of the particles involved in this thesis, e.g. hadrons such as protons or neutrons, are \vocab{composite}:
%
they are made up of fundamental pieces, and are not fundamental themselves.
%
Other particles, such as the quarks or gluons which make up the hadrons, are \vocab{fundamental}:
%
they cannot be written as a combination of more fundamental pieces.

\remark{}{
    In technical language, we often say that a particle is \textit{a representation of the Poincar\'e group}, and that a fundamental particle is \textit{an irreducible representation} (irrep).
}

\remark{}{
    The story does not end there, and there is a forest of subtleties which lie far beyond the scope of this thesis.
    %
    One important subtlety is that in \textit{gauge theories} such as \gls{qcd} and the \gls{sm}, even the description of fundamental particles as irreducible representations of the Poincar\'e group has flaws;
    %
    in technical language, gauge invariant states of a gauge theory are associated with additional gauge fields and asymptotic charge that are not encompassed by the definition of particles as irreps of the Poincar\'e group.
    %
    See e.g. the influential early work of \Reff{Kulish:1970ut}, which states that ``the representation of the Poincar\'e group ... acting in the physical Hilbert space ... does not contain discrete irreducible terms with nonvanishing mass. In other words, one can say that the relativistic concept of a charged particle does not exist.''
}

@@

In Yang-Mills theory with gauge group SU(N), we instead write \(U_{(R)} = e^{-i g T_{(R)}^a \alpha^a(x)}\) is upgraded to an element of (a representation of) SU(N), where the \(\{T^a_{(R)}\}\) are linearly-independent, traceless matrices:
%
the generators of SU(N) in the representation \(R\).
%
The index \(a\) takes values in \(a \in \{1,\,\cdots,\,N^2-1\}\), and in any representation the generators are guaranteed to satisfy the condition
\begin{align}
    \comm{T^a_{(R)}}{T^b_{(R)}}
    =
    i f^{abc} T^c_{(R)}
    \,,
\end{align}
for geometric reasons that we do not cover here.
%
The \(f^{abc}\) are numbers, and are called the \vocab{structure constants} of the group.
%
The generators are conventionally chosen to satisfy
\begin{align}
    \Tr\le(T_{(R)}^a T_{(R)}^b\ri) = \frac{1}{2} \delta^{ab}
    \,,
\end{align}
where the factor 1/2 is called the \textit{Dynkin index} (after mathematician Eugene Dynkin), and we have chosen it to be independent of representation.
%
These requirements on the generators can be used to show that the \(f^{abc}\) are completely antisymmetric with respect to exchange of any two indices.
%
When we have expressions which are representation independent, we will suppress the \((R)\) subscript, e.g. \(\Tr (T^a T^b) = \delta^{ab}/2\).

@@

Matter also matters.
%
Quarks in particular will play a large role in our discussion.
%
A \textit{matter field} \(\varphi\) -- e.g. a scalar or fermion field -- \vocab{transforms under the representation \(R\)} (of the gauge group) if
\begin{align}
    \varphi^\prime(x) = U_{(R)}(x) \, \varphi(x)
    \,.
\end{align}
%
A gauge theory with fermions \(\{\psi_{i}\}\) transforming in the representations \(R_i\) of the gauge group and scalars \(\{\phi_k\}\) transforming in the representations \(R_k\) can be described in terms of the \vocab{Yang-Mills-matter Lagrangian}%
\footnote{
    Non-Lagrangian methods for understanding \gls{qft}, such as bootstrap methods (e.g. the \textit{amplitudes program} of \gls{qft}) and the operator product expansion, are also available.
    %
    They can provide more rigorous descriptions of quantum dynamics than we pursue in this thesis, but are often more technically difficult as well.
}.
%
With massless Weyl fermions (guaranteed in the \gls{sm} due to the chiral representations of the fermionic \gls{sm} matter content), in \(d=3+1\) dimensions, and ignoring higher-dimension operators, it takes the form
\begin{subequations}
    \label{eq:yang-mills-matter-lagrangian}
\begin{align}
    \mathcal{L}
    &=
    \mathcal{L}_\text{YM}
    +
    \mathcal{L}_\text{fermion}
    +
    \mathcal{L}_\text{scalar}
    +
    \mathcal{L}_\text{Yukawa}
    \,,
    \\
    \mathcal{L}_\text{YM}
    &=
    -
    \frac{1}{2} \Tr F_{\mu\nu} F^{\mu\nu}
    -
    \frac{1}{8 \pi^2}
    \epsilon^{\mu\nu\rho\sigma}
    \Tr
        \Theta_\text{CP}
        F_{\mu\nu}
        F_{\rho\sigma}
    \,
    \\
    \mathcal{L}_\text{fermion}
    &=
    i \sum_i \psi_i^\dagger \overline{\sigma}^\mu D_{(R_i)\,\mu} \psi_i
    \\
    \mathcal{L}_\text{scalar}
    &=
    \sum_k \le(D_{(R_k)}^\mu \phi_k\ri)^\dagger D_{(R_k)\,\,\mu} \phi_k
    -
    \sum_{k,\ell} \! {}^\prime
    \,\,
    m_{k\ell}^2
    \,\,
    \phi_k \, \phi_\ell
    \,\,
    S_{R_k R_\ell}
    +
    \text{h.c}
    \\
    \notag
    &\qquad\qquad
    +
    \sum_{k,\ell,m,n} \!\!\!\! {}^\prime
    \,\,
    \lambda_{k\ell m n}
    \,\,
    \phi_k \, \phi_\ell \, \phi_m \, \phi_n
    \,\,
    S_{R_k R_\ell R_m R_n}
    +
    \text{h.c.}
    \\
    \mathcal{L}_\text{Yukawa}
    &=
    \sum_{ij}
    \sum_{k} \! {}^\prime
    \,\,
    y_{ijk}
    \,\,
    \psi_i \,\psi_j \,\, \phi_k
    \,\,
    S_{R_i R_j R_k}
    +
    \text{h.c.}
    \,,
\end{align}
\end{subequations}
where the \(\sum'\) indicates a sum on both the scalar fields \(\phi_k\) and the complex conjugates \(\phi_k^\dagger\), the \(S\) are invariant symbols of the gauge group which project onto singlet representations and guarantee the invariance of the Lagrangian under gauge transformations, the CP-violating angle \(\Theta_\text{CP}\) -- usually multiplied by an additional factor of \(g^2\) -- is a (hermitian) central charge of the gauge group (represented in the adjoint representation), and ``h.c.'' means hermitian conjugate.

\remark{}{
    A single degree of freedom of the central charge \(\Theta_\text{CP}\) can be removed for every anomalous global symmetry of \(\mathcal{L}\), and CP-violating angles associated with Abelian (U(1) or \(\mathbb{R}\)) factors of the gauge group can be ignored in topologically trivial spaces.
    %
    See, for example, \Reff{Tong:2017oea}.
}

@@

In this thesis, we will only be using the rules involving three-point vertices and external lines.
%
Other factors that will make consistent appearances in the discussions of this thesis are the \vocab{quadratic Casimirs} \(C_R\) and Dynkin indices \(T_R\) -- called \textit{group theory factors} -- defined by
\begin{align}
    \sum_a
    T^a_{(R)} T^a_{(R)}
    =:
    C_R
    1\!\!1_{\,(R)}
    \,,
\end{align}
where \(1\!\!1_{\,(R)}\) indicates the identity matrix in the representation \(R\), and
\begin{align}
    \Tr T^a_{(R)} T^b_{(R)}
    =:
    T_R \delta^{ab}
    \,.
\end{align}
In an unfortunate similarity of notation, the Dynkin index of the representation \(R\) is also denoted by a capital \(T\);
%
luckily, our notation is nonetheless unambiguous, as the Dynkin index \(T_R\) will never appear with a color index \(a\).

For the purposes of this thesis, we will use the standard normalization of the generators of SU(N) to obtain
\begin{align}
    C_A = N_c
    \qquad
    C_R = \frac{N_c^2 - 1}{2 N_c}
    \qquad
    T_F = \frac{1}{2}
    \,,
\end{align}
where \(A\) denotes the adjoint representation, under which gluons transform, and \(F\) denotes the fundamental representation, under which quarks transform.

@@

\begin{exercise}
    \label{ex:lqcd-nonpert}
    Using the \gls{qcd} beta function, find the energy scale \(\mu = \LambdaQCD\) at which the strong coupling becomes infinite, \(\alpha_s(\LambdaQCD) := \infty\), in terms of the coupling \(\alpha_s := \alpha_s(\mu_0)\) given at a fixed initial scale \(\mu_0\).
    %
    Prove that \(\LambdaQCD\) it cannot be expressed in terms of a power series in \(\alpha_s\), and that power corrections of the form \((\LambdaQCD/Q)^n\) are indeed non-perturbative.
\end{exercise}

@@

\remark{}{
    The Fadeev-Kulish (FK) formalism for dressed, gauge invariant scattering \cite{Kulish:1970ut}, inspired by the preliminary work of P.A.M. Dirac \cite{Dirac:1955uv} and the later work of \Reff{Chung:1965zza}, offers a complementary perspective on IR finiteness (see also \Reff{Contopanagos:1991yb} for more developments on the KLN perspective).
    %
    The FK formalism operates under the assumption that infra-red finiteness may be ensured by working with appropriately defined initial and final states involving \vocab{dressed} charged particles -- roughly, particles surrounded by clouds of gauge bosons.
    %
    See \Reffs{Catani:1984dp,Ciafaloni:1989vs,Lavelle:1995ty,Bagan:1999jf,Larkoski:2014bxa,Frye:2018xjj,Hannesdottir:2021oqb} for additional perspectives and discussions on dressed states in \gls{qcd}.
}

@@

\begin{figure}[t!]
    \centering
    \includegraphics[width=0.7\textwidth]{example-image-a}

    % Caption
    \caption[A visualization of \gls{preconfinement}\sam{incomplete}]{
        A visualization of \gls{preconfinement}, a local manifestation of \glslink{phd}{parton-hadron duality} in which partons tend to form low-mass color-singlet clusters. \sam{complete}
    }
    \label{fig:preconfinement}
\end{figure}

@@

\remark{}{
    A \textbf{Warning!} is warranted at this stage:
}
\vspace{-20pt}

@@

\remark{ignored-effects}{
    While we also averaged over final state spin information, we we did not \emph{have} to do so, and we can also define analogous \textit{polarized} correlation matrices which depend also on the spin (or other quantum numbers) of the outgoing states \(i\) and \(j\).
    %
    On the other hand, there are several important effects that are thrown out by the stringent assumptions of \Prop{correlation-matrix-factorization} that cannot be so easily recovered.
    %
    Notably, the strict collinear limit we consider in this thesis erases additional kinematic and color correlations associated with \emph{spectators} of the splitting \cite{Catani:1996vz,Hoche:2014rga}.
}

@@

Traditionally, one states that virtual contributions must come in to save the day and regularize the splitting functions because of the KLN theorem (\Thm{kln}).
%
But I have never seen a proof or an argument that has fully satisfied me, and I have been hurt too many times (see \Sec{irc-safety}, especially Lore~\ref{lore:irc-safety-lore} and Remark~\ref{rem:multiple-soft-emissions-break-invariance}, and e.g. my first relationship).
%
\textbf{Within the framework I have derived above, I know of no way to systematically describe the regularization of the splitting functions we found above and thus to derive their probabilistic interpretation.}

@@

\begin{figure}[]
    \centering
    \includegraphics[width=0.6\textwidth]{example-image-a}
    \caption[A cartoon depicting charge coherence in QED. \sam{incomplete}]{%
        \sam{ Cartoon depicting two-particle examples of charge coherence in QED.  }
    }
    \label{fig:qed-coherent-branching}
\end{figure}

@@

\begin{figure}[]
    \centering
    \includegraphics[width=0.8\textwidth]{example-image-a}
    \caption[A cartoon depicting color coherence and angular ordering in QCD. \sam{incomplete}]
    {
        \sam{ Cartoons depicting angular ordering as a coherence effect in QCD.  %
            Qualitative analogy, imagining gluon radiation, analogized as ripples in a pool of water, produced by charged \gls{qcd} matter, analogized as rocks dropped into the water which source the ripples.
        }
    }
    \label{fig:qcd-coherent-branching}
\end{figure}

@@

@@

\begin{exercise}
    \label{ex:qcd-coherent-branching-proof}
    Don't get intimidated by the Lemma below.
    %
    Instead, by following the fixed-order logic of \Lem{collinear-color-coherence} and \Example{charge-coherence}, and the recursive logic of \Exercise{charge-coherence-mparticle} in the limit of a hierarchy of collinear splittings, demonstrate the phenomenon of \vocab{QCD Coherent Branching}:
\end{exercise}

@@

\sam{correct group theory?}
\begin{align}
    T_A = \frac{1}{2}
    \,.
\end{align}

@@

In fact,
\begin{align}
    C_R = \frac{T_A C_A}{T_R}
    \,,
\end{align}
\sam{correct group theory?}
which can be quickly and amusingly shown with \vocab{bird track diagrams} \cite{Cvitanovic:2008zz,Keppeler:2017kwt}.

@@

\subsection{Splitting Functions}

Finally, we tabulate the properties of the exclusive and inclusive \gls{qcd} \glspl{splittingfn} in \Tabs{exclusive-splitting}{inclusive-splitting} respectively.
%
We use \(B(m,n)\) to denote Euler's beta function,
\begin{align}
    B(m,n) = \frac{\Gamma(m)\Gamma(n)}{\Gamma(m+n)}
    \,.
\end{align}

\sam{Missing delta functions/virtual contributions}
%

% +===================================
\begin{table}[h!]
\begin{center}
\LARGE $\boldsymbol{P_{ij \leftarrow \ell}(z)}$
\end{center}
\vspace{-10pt}
\centering
\caption[Exclusive NLO QCD splitting functions and their Mellin moments.]{
    Exclusive QCD splitting functions and their Mellin moments at NLO (\(z = E_i/E_\ell\)).
    %
    $B(m,n)$ denotes Euler's beta function.  \sam{check}
    %
    \sam{include comment about \(N_f\)}
}
\label{tab:exclusive-splitting}
\vspace{5pt}
\renewcommand{\arraystretch}{3.2}
\rowcolors{1}{gray!20}{gray!5}
\begin{tabular}{|>{\bfseries}m{2cm}|m{4cm}|m{7cm}|}
\hline
\centering \textbf{Channel} & \centering \textbf{Splitting Function} &
\centering $\boldsymbol{\hat{P}(m,n)} := \int_0^1 \dd x
\, x^{m-1} \le(1-x\ri)^{n-1} \, \, P(x)$
\tabularnewline
\hline
\centering $gq \leftarrow q$
                            &
\centering $C_F \dfrac{1 + (1 - z)^2}{z}$
                            &
\centering
$C_F \le(B\le(m,n-1\ri) + B\le(m+2,n-1\ri)\ri)$
\tabularnewline
\hline
\centering
$gg \leftarrow g$
                            &
\centering $2 C_A \dfrac{(1 - z(1 - z))^2}{z(1 - z)}$
                            &
\centering
$2 C_A \big(B\le(m-1, n+1\ri) + B\le(m+1, n-1\ri)$
\centering
$+ B\le(m+1, n+1\ri)\big)$
\tabularnewline
\hline
\centering $q\bar{q} \leftarrow g$
                            &
\centering $T_F \le(z^2 + (1-z)^2\ri)$
                            &
\centering
$T_F \le(B\le(m+2, n\ri) + B\le(m, n+2\ri)\ri)$
\tabularnewline
\hline
\end{tabular}
\end{table}
% +===================================

>>>>

Rather than scattering rocks, we will be concerned with the scattering of subatomic particles.
%
But what is a particle?
%
To answer this, let us point out that an electron moving to your left is a different state than an electron moving to your right;
%
furthermore, if you say that it is moving to the left, someone facing you will say that it is moving to the right!
%
But you should both agree, regardless of the direction of its motion or spin, that it is the same \textit{type} of particle:
%
an electron.
%
So a particle should be a \textit{set} or \textit{collection} of states which different (inertial) observers agree correspond to the same particle.

\begin{definitionbox}{Particle}{particle}
    A \vocab{particle} is a \textit{set} of quantum states characterized by mass, spin, and charge -- their \textit{quantum numbers} -- such that:
    \begin{itemize}
        \item
            it is \vocab{invariant} under Poincar\'e transformations:
            %
            the overall set of states remains unchanged under translations in space/time, rotations, and boosts;

        \item
            It is \vocab{irreducible}:
            %
            it contains no proper invariant subsets.
    \end{itemize}
    In technical language, we often say that a particle is \textit{an irreducible representation} (irrep).
\end{definitionbox}

\remark{}{
    Really, this definition holds in flat spacetimes when using only massive particles.
    %
    The story does not end there, however, and beyond there is a forest of subtleties which lie far beyond the scope of this thesis.
    %
    One important subtlety is that in \textit{gauge theories} such as \gls{qcd} and the \gls{sm}, even the description of fundamental particles as irreducible representations of the Poincar\'e group has flaws;
    %
    in technical language, gauge invariant states of a gauge theory are associated with additional gauge fields and asymptotic charge that are not encompassed by the definition of particles as irreps of the Poincar\'e group.
    %
    See e.g. the influential early work of \Reff{Kulish:1970ut}, which states that ``the representation of the Poincar\'e group ... acting in the physical Hilbert space ... does not contain discrete irreducible terms with nonvanishing mass. In other words, one can say that the relativistic concept of a charged particle does not exist.''
    %
    The idea that the notion of ``particle'' must be modified in theories with massless particles manifests in several frameworks.
    %
    The framework closest to the ideas presented in this work (see especially \Sec{splitting-functions}) states that we must stop thinking about quantum amplitudes and begin thinking about \emph{inclusive probabilities} \cite{Yennie:1961ad}.
    %
    This perspective can also be connected to the framework of \emph{infraparticles} \cite{Schroer:1963gw,Schroer:2008gd} (also the Faddeev-Kulish framework \cite{Dirac:1955uv,Kulish:1970ut}).
}

@@

In Yang-Mills theory with gauge group SU(N), we instead write \(U_{(R)} = e^{-i g T_{(R)}^a \alpha^a(x)}\) is upgraded to an element of (a representation of) SU(N), where the \(\{T^a_{(R)}\}\) are linearly-independent, traceless matrices:
%
the generators of SU(N) in the representation \(R\).
%
The index \(a\) takes values in \(a \in \{1,\,\cdots,\,N^2-1\}\), and in any representation the generators are guaranteed to satisfy the condition
\begin{align}
    \comm{T^a_{(R)}}{T^b_{(R)}}
    =
    i f^{abc} T^c_{(R)}
    \,,
\end{align}
for geometric reasons that we do not cover here.
%
The \(f^{abc}\) are numbers, and are called the \vocab{structure constants} of the group.
%
The generators are conventionally chosen to satisfy
\begin{align}
    \Tr\le(T_{(R)}^a T_{(R)}^b\ri) = T_R \delta^{ab}
    \,,
\end{align}
where the factor 1/2 is called the \textit{Dynkin index} (after mathematician Eugene Dynkin), which we will discuss more when we introduce the Feynman rules of \gls{qcd}.
%
These requirements on the generators can be used to show that the \(f^{abc}\) are completely antisymmetric with respect to exchange of any two indices.

@@

Matter also matters.
%
Quarks in particular will play a large role in our discussion.
%
A \textit{matter field} \(\varphi\) -- e.g. a scalar or fermion field -- \vocab{transforms under the representation \(R\)} (of the gauge group) if
\begin{align}
    \varphi^\prime(x) = U_{(R)}(x) \, \varphi(x)
    \,.
\end{align}
%
A gauge theory with fermions \(\{\psi_{i}\}\) transforming in the representations \(R_i\) of the gauge group and scalars \(\{\phi_k\}\) transforming in the representations \(R_k\) can be described in terms of the \vocab{Yang-Mills-matter Lagrangian}%
\footnote{
    Non-Lagrangian methods for understanding \gls{qft}, such as bootstrap methods (e.g. the \textit{amplitudes program} of \gls{qft}) and the operator product expansion, are also available.
    %
    They can provide more rigorous descriptions of quantum dynamics than we pursue in this thesis, but are often more technically difficult as well.
}.
%
With massless Weyl fermions (guaranteed in the \gls{sm} due to the chiral representations of the fermionic \gls{sm} matter content), in \(d=3+1\) dimensions, and ignoring higher-dimension operators, it takes the form
\begin{subequations}
    \label{eq:yang-mills-matter-lagrangian}
\begin{align}
    \mathcal{L}
    &=
    \mathcal{L}_\text{YM}
    +
    \mathcal{L}_\text{fermion}
    +
    \mathcal{L}_\text{scalar}
    +
    \mathcal{L}_\text{Yukawa}
    \,,
    \\
    \mathcal{L}_\text{YM}
    &=
    -
    \frac{1}{2} \Tr F_{\mu\nu} F^{\mu\nu}
    -
    \frac{1}{8 \pi^2}
    \epsilon^{\mu\nu\rho\sigma}
    \Tr
        \Theta_\text{CP}
        F_{\mu\nu}
        F_{\rho\sigma}
    \,
    \\
    \mathcal{L}_\text{fermion}
    &=
    i \sum_i \psi_i^\dagger \overline{\sigma}^\mu D_{(R_i)\,\mu} \psi_i
    \\
    \mathcal{L}_\text{scalar}
    &=
    \sum_k \le(D_{(R_k)}^\mu \phi_k\ri)^\dagger D_{(R_k)\,\,\mu} \phi_k
    -
    \sum_{k,\ell} \! {}^\prime
    \,\,
    m_{k\ell}^2
    \,\,
    \phi_k \, \phi_\ell
    \,\,
    S_{R_k R_\ell}
    +
    \text{h.c}
    \\
    \notag
    &\qquad\qquad
    +
    \sum_{k,\ell,m,n} \!\!\!\! {}^\prime
    \,\,
    \lambda_{k\ell m n}
    \,\,
    \phi_k \, \phi_\ell \, \phi_m \, \phi_n
    \,\,
    S_{R_k R_\ell R_m R_n}
    +
    \text{h.c.}
    \\
    \mathcal{L}_\text{Yukawa}
    &=
    \sum_{ij}
    \sum_{k} \! {}^\prime
    \,\,
    y_{ijk}
    \,\,
    \psi_i \,\psi_j \,\, \phi_k
    \,\,
    S_{R_i R_j R_k}
    +
    \text{h.c.}
    \,,
\end{align}
\end{subequations}
where the \(\sum'\) indicates a sum on both the scalar fields \(\phi_k\) and the complex conjugates \(\phi_k^\dagger\), the \(S\) are invariant symbols of the gauge group which project onto singlet representations and guarantee the invariance of the Lagrangian under gauge transformations, the charge-parity symmetry violating (CP-violating) parameter \(\Theta_\text{CP}\) -- usually multiplied by an additional factor of \(g^2\) -- is a central element of the gauge Lie algebra (almost -- see below) in the adjoint representation, and ``h.c.'' means hermitian conjugate.

\remark{}{
    The CP-violating terms in a gauge theory are usually denoted by a set of real parameters \(\{\theta_i\}\), one for each simple factor \(G_i\) of the gauge group \(G = \prod_i G_i\).
    %
    Each \(\theta_i\) multiplies the topological density \(\Tr(F^{(i)}_{\mu\nu} \tilde{F}^{(i)\,\mu\nu})\), associated with the field strength \(F^{(i)}\) of \(G_i\), and breaks CP symmetry when nonzero.
    %
    In the Lagrangian of \Eq{yang-mills-matter-lagrangian}, these terms are collectively encoded by \(\Theta_\text{CP}\), a single object in the center of the \emph{universal enveloping algebra} of the Lie algebra of the gauge group.%
    \footnote{
        The Lie algebra alone does not work.
        %
        For example, in SU(\(N\)), the adjoint representation of the algebra only contains traceless elements, while the CP-violating angle \(\theta_\text{CP}\) should be proportional to the identity to yield standard definitions.
        %
        Using the universal enveloping algebra in the adjoint representation allows for terms that are proportional to the identity, and thus the standard CP-violating angles \(\theta_\text{CP}\).
    }
    %
    Furthermore, a single degree of freedom of \(\Theta_\text{CP}\) can be removed for every anomalous global symmetry of \(\mathcal{L}\), and CP-violating angles associated with Abelian (U(1) or \(\mathbb{R}\)) factors of the gauge group can be ignored in topologically trivial spaces (see \Reff{Tong:2017oea} for a pedagogical introduction with an emphasis on the \gls{sm}).
}

@@

In this thesis, we will only be using the rules involving three-point vertices and external lines.
%
Other factors that will make consistent appearances in the discussions of this thesis are the \vocab{quadratic Casimirs} \(C_R\) and Dynkin indices \(T_R\) -- called \textit{group theory factors} -- defined by
\begin{align}
    \sum_a
    T^a_{(R)} T^a_{(R)}
    =:
    C_R
    1\!\!1_{\,(R)}
    \,,
\end{align}
where \(1\!\!1_{\,(R)}\) indicates the identity matrix in the representation \(R\), and
\begin{align}
    \Tr T^a_{(R)} T^b_{(R)}
    =:
    T_R \, \delta^{ab}
    \,.
\end{align}
In an unfortunate similarity of notation, the Dynkin index of the representation \(R\) is also denoted by a capital \(T\);
%
luckily, our notation is nonetheless unambiguous, as the Dynkin index \(T_R\) will never appear with a color index \(a\).

For the purposes of this thesis, we will use the standard normalization of the generators of SU(N) to obtain
\begin{align}
    C_A = N_c
    \qquad
    C_F = \frac{N_c^2 - 1}{2 N_c}
    \qquad
    T_F = \frac{1}{2}
    \,,
\end{align}
where \(A\) denotes the adjoint representation, under which gluons transform, and \(F\) denotes the fundamental representation, under which quarks transform.

\begin{exercise}
    \label{ex:group-theory-factors}
    Show that the group theory factors for the representation \(R\) are given by
    \begin{align}
        T_R \, \text{dim}(A)
        =
        C_R \, \text{dim}(R)
        \,,
    \end{align}
    where \(\text{dim}(R)\) indicates the dimension of the representation \(R\), and \(\text{dim}(A)\) -- the number of generators and the dimension of the adjoint representation -- is \(N_c^2 - 1\) for SU(\(N_c\)).

    What is the Dynkin index \(T_A\) of the adjoint representation?
\end{exercise}

@@

\begin{exercise}
    \label{ex:lqcd-nonpert}
    Using the \gls{qcd} beta function, find the energy scale \(\mu = \LambdaQCD\) at which the strong coupling becomes infinite, \(\alpha_s(\LambdaQCD) := \infty\), in terms of the coupling \(\alpha_s := \alpha_s(\mu_0)\) given at a fixed initial scale \(\mu_0\).
    %
    Prove that \(\LambdaQCD\) cannot be expressed in terms of a power series in \(\alpha_s\), and that power corrections of the form \((\LambdaQCD/Q)^n\) are indeed non-perturbative.
\end{exercise}

@@

\remark{faddeev-kulish}{
    The Faddeev-Kulish (FK) formalism for dressed, gauge invariant scattering \cite{Kulish:1970ut}, inspired by the preliminary work of P.A.M. Dirac \cite{Dirac:1955uv} and the later work of \Reff{Chung:1965zza}, offers a complementary perspective on IR finiteness (see also \Reff{Contopanagos:1991yb} for more developments on the KLN perspective).
    %
    The FK formalism operates under the assumption that infra-red finiteness may be ensured by working with appropriately defined initial and final states involving \vocab{dressed} charged particles -- roughly, particles surrounded by clouds of gauge bosons.
    %
    See \Reffs{Catani:1984dp,Ciafaloni:1989vs,Lavelle:1995ty,Bagan:1999jf,Larkoski:2014bxa,Frye:2018xjj,Hannesdottir:2021oqb} for additional perspectives and discussions on dressed states in \gls{qcd}.
}

@@

@@

\remark{ward-warning}{
    A \textbf{Warning!} is warranted at this stage:
}
\vspace{-20pt}

@@

\remark{ignored-effects}{
    While we also averaged over final state spin information, we did not \emph{have} to do so, and we can also define analogous \textit{polarized} correlation matrices which depend also on the spin (or other quantum numbers) of the outgoing states \(i\) and \(j\).
    %
    On the other hand, there are several important effects that are thrown out by the stringent assumptions of \Prop{correlation-matrix-factorization} that cannot be so easily recovered.
    %
    Notably, the strict collinear limit we consider in this thesis erases additional kinematic and color correlations associated with \emph{spectators} of the splitting \cite{Catani:1996vz,Hoche:2014rga}.
}

@@

Traditionally, one states that virtual contributions must come in to save the day and regularize the splitting functions because of the KLN theorem (\Thm{kln}).
%
But I have never seen a proof or an argument that has fully satisfied me, and I have been hurt too many times (see \Sec{irc-safety} -- especially Lore~\ref{lore:irc-safety-lore} and Remark~\ref{rem:multiple-soft-emissions-break-invariance} -- and Remark~\ref{rem:ward-warning}).
%
Therefore I warn that

\remark{i-dont-know-splitting}{
    \textbf{Within the framework I have derived above, I know of no way to regularize our splitting functions or derive their probabilistic interpretation.}
}

@@

\remark{}{
    For a visualization of charge coherence in the non-Abelian, more complicated context of \gls{qcd}, see \Fig{qcd-coherent-branching}.
}

@@

@@

\begin{figure}[t!]
    \centering
    \includegraphics[width=1.05\textwidth]{figures/misc/color_coherence.pdf}
    \caption[A diagram illustrating color coherence and angular ordering in QCD.]
    {
        A diagram illustrating the first stage of color coherence and angular ordering in \gls{qcd}, whose recursive application leads to \Lem{qcd-coherent-branching}.
        %
        Near either individual parton \(i\in\{1,2\}\), soft gluons are emitted with a probability proportional to the \(T_i^2 = C_{R_i}\);
        %
        the overlap in which these probabilities add is striped red and blue.
        %
        However, in the region surrounding \emph{both} partons, indicated in purple, their charges add coherently and soft gluons are emitted with a probability proportional to \(C_{R_1\oplus R_2}\).
        %
        Beyond the collinear limit, a factorization-breaking \emph{fringe} (indicated as rainbow-colored and dotted) emerges around the coherent region, where it is less meaningful to talk about the probability of emitting a soft gluon;
        %
        in this region, the total differential cross section \(\dd\sigma_{N+1}\) \emph{does not factorize} into the product of an \(N\) particle differential cross section \(\dd\sigma_{N}\) and a soft gluon emission probability.
    }
    \label{fig:qcd-coherent-branching}
\end{figure}

@@

\begin{exercise}
    \label{ex:qcd-coherent-branching-proof}
    Don't get intimidated by the Lemma below.
    %
    Instead, by following the fixed-order logic of \Lem{collinear-color-coherence} and \Example{charge-coherence}, and the recursive logic of \Exercise{charge-coherence-mparticle} in the limit of a hierarchy of collinear splittings, demonstrate the phenomenon of \vocab{QCD Coherent Branching} (see also \Fig{qcd-coherent-branching}):
\end{exercise}

@@

@@

In fact, we showed in \Exercise{group-theory-factors} that
\begin{align}
    C_R = \frac{T_A C_A}{T_R}
    \,,
\end{align}
which can be quickly shown with \vocab{bird track diagrams} \cite{Cvitanovic:2008zz,Keppeler:2017kwt}.

@@

%\subsection{Splitting Functions}

%Finally, we tabulate the properties of the exclusive and inclusive \gls{qcd} \glspl{splittingfn} in \Tabs{exclusive-splitting}{inclusive-splitting} respectively.
%%
%We use \(B(m,n)\) to denote Euler's beta function,
%\begin{align}
%    B(m,n) = \frac{\Gamma(m)\Gamma(n)}{\Gamma(m+n)}
%    \,.
%\end{align}

%\sam{Missing delta functions/virtual contributions}
%%

%% +===================================
%\begin{table}[h!]
%\begin{center}
%\LARGE $\boldsymbol{P_{ij \leftarrow \ell}(z)}$
%\end{center}
%\vspace{-10pt}
%\centering
%\caption[Exclusive NLO QCD splitting functions and their Mellin moments.]{
%    Exclusive QCD splitting functions and their Mellin moments at NLO (\(z = E_i/E_\ell\)).
%    %
%    $B(m,n)$ denotes Euler's beta function.  \sam{check}
%    %
%    \sam{include comment about \(N_f\)}
%}
%\label{tab:exclusive-splitting}
%\vspace{5pt}
%\renewcommand{\arraystretch}{3.2}
%\rowcolors{1}{gray!20}{gray!5}
%\begin{tabular}{|>{\bfseries}m{2cm}|m{4cm}|m{7cm}|}
%\hline
%\centering \textbf{Channel} & \centering \textbf{Splitting Function} &
%\centering $\boldsymbol{\hat{P}(m,n)} := \int_0^1 \dd x
%\, x^{m-1} \le(1-x\ri)^{n-1} \, \, P(x)$
%\tabularnewline
%\hline
%\centering $gq \leftarrow q$
%                            &
%\centering $C_F \dfrac{1 + (1 - z)^2}{z}$
%                            &
%\centering
%$C_F \le(B\le(m,n-1\ri) + B\le(m+2,n-1\ri)\ri)$
%\tabularnewline
%\hline
%\centering
%$gg \leftarrow g$
%                            &
%\centering $2 C_A \dfrac{(1 - z(1 - z))^2}{z(1 - z)}$
%                            &
%\centering
%$2 C_A \big(B\le(m-1, n+1\ri) + B\le(m+1, n-1\ri)$
%\centering
%$+ B\le(m+1, n+1\ri)\big)$
%\tabularnewline
%\hline
%\centering $q\bar{q} \leftarrow g$
%                            &
%\centering $T_F \le(z^2 + (1-z)^2\ri)$
%                            &
%\centering
%$T_F \le(B\le(m+2, n\ri) + B\le(m, n+2\ri)\ri)$
%\tabularnewline
%\hline
%\end{tabular}
%\end{table}
%% +===================================
